---
title       : "R 기초과정 02"
subtitle    : "집단비교"
author      : "김형준"
job         : "Analytic Director / (주) 퀀트랩 / kim@mindscale.kr "
logo        : logo_03.png
license     : by-nc-sa
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
widgets     : []            # {mathjax, bootstrap, quiz}
mode        : selfcontained
hitheme     : tomorrow      # {tomorrow, tomorrow_night, solarized_dark, solarized_light}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
knit        : slidify::knit2slides

---

## 기초 2 과정

### 학습 목표

패키지 설치하기  
데이터 편집하기  
Missing 다루기  
집단 평균 비교하기  
상관관계 이해하기

---  .new-background

## 패키지(라이브러리)
```{r, pac}
head(installed.packages())
```

---  .new-background

### 패키지 설치
```{r, i_pac, eval=T}
install.packages("bootstrap")
```

```{r, ii_pac, eval=F}
install.packages("bootstrap", repos="http://cran.rstudio.com/")
```

### 패키지 정보  

[CRAN](https://www.google.co.kr/webhp?sourceid=chrome-instant&ion=1&espv=2&ie=UTF-8#q=cran)  

[CRAN Pac. by names](http://cran.r-project.org/web/packages/available_packages_by_name.html)  
  
[CRAN Pac. by Tasks](http://cran.r-project.org/web/views/)  

---  .new-background

## 오늘의 데이터

```{r, echo=F}
par(family="NanumGothic")
setwd("/Users/kimhyungjun/Dropbox/repo/r_basic/html")
df <- read.csv("data/salesC.csv", fileEncoding = "CP949", stringsAsFactors = F)
str(df)
```

---  .new-background

```{r}
df <- read.csv("data/salesC.csv", fileEncoding = "CP949", stringsAsFactors = F, na.strings = c(".", "#"))
str(df)
```

---  .new-background

```{r}
table(df$성별)
table(df$요일)
table(df$점포)
table(df$매출월)
```

---  .new-background

### first vs latter

```{r}
df.first <- subset(df, 매출월 <= 6 )
mean(df.first$구매건수)
mean(df.first$구매건수, na.rm = T)
df.latter <- subset(df, 매출월 > 6 )
mean(df.latter$구매건수, na.rm = T)
```

---  .new-background

### Histogram

```{r}
hist(df.first$구매건수, xlab = "구매건수", main = "first")
```

---  .new-background

### Histogram

```{r}
hist(df.latter$구매건수, xlab = "구매건수", main = "latter")
```

---  .new-background

### first vs latter

```{r, message=F}
library(bootstrap)
```

```{r}
first <- na.omit(df.first$구매건수)
latter <- na.omit(df.latter$구매건수)
first.means <- bootstrap(first, 5, mean)$thetastar
latter.means <- bootstrap(latter, 5, mean)$thetastar
```

---  .new-background .modal

hist(first.means - latter.means , main = "상반기 - 하반기", xlab = "평균 차이")

---  .new-background .modal

### first vs latter

```{r}
first.means <- bootstrap(first, 10, mean)$thetastar
latter.means <- bootstrap(latter, 10, mean)$thetastar
hist(first.means - latter.means, main = "상반기 - 하반기", xlab = "평균 차이")
```

---  .new-background .modal

### first vs latter

```{r}
first.means <- bootstrap(first, 15, mean)$thetastar
latter.means <- bootstrap(latter, 15, mean)$thetastar
hist(first.means - latter.means, main = "상반기 - 하반기", xlab = "평균 차이")
```

---  .new-background .modal

### first vs latter

```{r}
first.means <- bootstrap(first, 5000, mean)$thetastar
latter.means <- bootstrap(latter, 5000, mean)$thetastar
hist(first.means - latter.means, main = "상반기 - 하반기", xlab = "평균 차이")
```

---  .new-background .modal

### first vs first

```{r}
hist(first.means - first.means, main = "상반기 - 상반기", xlab = "평균 차이")
```

--- &twocol w1:50% w2:50% .new-background

*** =left

```{r}
hist(first.means - latter.means, main = "상반기 - 하반기", xlab = "평균 차이")
```

*** =right

```{r}
hist(first.means - first.means, main = "상반기 - 상반기", xlab = "평균 차이")
```


---  .new-background

```{r}
var.test(df.first$구매건수, df.latter$구매건수)
```

---  .new-background

```{r}
t.test(df.first$구매건수, df.latter$구매건수)
```

---  .new-background

```{r}
df.sunday <- subset(df, 요일 ==  "일")
```

```{r}
var.test(구매건수 ~ 성별,  df.sunday)
```

---  .new-background

```{r}
t.test(구매건수 ~ 성별, df.sunday)
```

---  .new-background

```{r}
df.holiday <- subset(df, 요일 %in% c("토", "일"))
```

```{r}
var.test(구매건수 ~ 성별,  df.holiday)
```

---  .new-background

```{r}
t.test(구매건수 ~ 성별, df.holiday)
```

---  .new-background

```{r}
x = 1:10; y = c(1,3,2,4,5,6,7,9,8,10);
plot(x, y)
cor(x,y)
```

---  .new-background

```{r}
x = 1:10; y = c(5,10,1,6,4,3,7,9,8,3);
plot(x, y)
cor(x,y)
```

---  .new-background

```{r}
df.sex <- split(df, df$성별)
str(df.sex)
str(df.sex['남'])
```

---  .new-background

```{r}
x <- data.frame(sex = "남자", value1 = 123, value2 = 125)
y <- data.frame(sex = "여자", value1 = 122)
rbind(x,x)
#rbind(x,y)
list(x,y)
```

---  .new-background .modal

```{r}
df.high <- df[df$구매건수 >= quantile(df$구매건수, probs = .9, na.rm = T), ]
mosaicplot(df.high$성별 ~ df.high$요일, 
           main="성별 by 요일", shade=F, 
           color = 1:7, xlab="성별", ylab="요일")
```

---  .new-background .modal

```{r}
mosaicplot(df.high$성별 ~ factor(df.high$요일, levels=c("월", "화", "수", "목", "금", "토", "일")), 
           main="성별 by 요일", shade=F, 
           color = 1:7, xlab="성별", ylab="요일")
```

---  .new-background .modal

```{r}
df$요일 <- factor(df$요일, levels = c("월", "화", "수", "목", "금", "토", "일"))
df.month <- sapply(split(df, df$요일), function(x) x$구매건수)
cor(df.month, use = "pairwise")
```

---  .new-background .modal

```{r}
cor(df.month, use = "complete")
```

